Report

### 2022-06-27 manuscript.Rmd
minor changes

Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`). `postnumb` is a continuous variable which ranges from 0-63. `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking. In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  site`:

@@

#specify Hypotheses to evaluate
hypotheses <- "site2=site1=site3=site4=site5; site2&site3>site1&site4&site5"
hypotheses <- "site2=site1=site3=site4=site5"

@@

The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually. The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low. When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually. For sample 3 the specified hypotheses are even more likely than their complement. This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation. If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation. This leads to a more extreme combined evaluation than any individual evaluation. Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.

### using bain objects
The `pbf()` function also works when objects are first converted to `bain` objects. In this specific example, each `bain` object corresponding to one of the three samples has the $H_0$ and $H_A$ in common, but contains one sample-specific hypothesis. The `pbf()` function only returns output for the hypotheses that are present in all results. When `bain` objects are passed to the `pbf()`, no additional hypotheses need to be passed to `pbf()` as they are already saved in the objects.

@@

When there is no hypothesis that is shared by each sample, `pbf()` throws an error.
```{r}
bains_different_hyps <- list(
  bain(lm_sample1, "site2=site1=site3=site4=site5"),
  bain(lm_sample2, "site2=site1=site3>site4=site5"),
  bain(lm_sample3, "site2&site3>site1&site4&site5")
)

@@

### using named vectors
Another specific use-case is when the raw data from different samples is not available. In this case, the researcher only has access to the parameter-values that are provided in the research papers. Because the `pbf()` function inherently converts objects to `bain` objects, each sample requires three pieces of information in order to be usable: the named parameter estimates, a covariance matrix of their standard errors and the sample size. First these values are extracted from all samples. A function is created that extracts is information which is then used to extract the necessary information from all samples.

```{r}
get_estimates <- function(sample){
  number_sites <- length(levels(sesamesim$site))

  sample_size <- nrow(sample)

  site_estimates <- sapply(1:number_sites, function(i){
    rbind(
      mean(sample[sample['site'] == i, "postnumb"]),
      sd(sample[sample['site'] == i, "postnumb"])
      )
    })

  est_names <- sprintf("site%d", 1:number_sites)
  site_means <- site_estimates[1,]
  names(site_means) <- est_names

  covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
  dimnames(covmat) <- list(est_names, est_names)
  return(list(means = site_means, covmat = covmat, sample_size = sample_size))
}

@@

Then, each object is converted to a `bain` object with these estimates
```{r}
bain_sample1_ests <- bain(x = ests_sample1[['means']], 
                          Sigma = ests_sample1[['covmat']], 
                          n = ests_sample1[['sample_size']], 
                          hypothesis = hypotheses)

>>>>

Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`). `postnumb` is a continuous variable which ranges from 0-63. `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking. In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  `site`:

@@

#specify Hypotheses to evaluate
hypotheses <- "site2=site1=site3=site4=site5; site2&site3>site1&site4&site5"

@@

The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually. The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low. When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually. For sample 3 the specified hypotheses are even more likely than their complement. This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation. If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation. This leads to a more extreme combined evaluation than any individual evaluation. Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true and PBF should be reflective of this.

### using bain objects
The `pbf()` function also works when objects are first converted to `bain` objects. In this specific example, each `bain` object corresponding to one of the three samples has the $H_0$ and $H_A$ in common, but contains one sample-specific hypothesis. The `pbf()` function only returns output for the hypotheses that are present in all results. When `bain` objects are passed to the `pbf()`, no additional hypotheses need to be passed to `pbf()` as they are already saved in the `bain` objects.

@@

When there is no hypothesis that is shared by each sample, `pbf()` throws an error.
```{r}
bains_different_hyps <- list(
  bain(lm_sample1, "site2=site1=site3=site4=site5"),
  bain(lm_sample2, "site2=site1=site3>site4=site5"),
  bain(lm_sample3, "site2&site3<site1&site4&site5")
)

@@

### using named vectors
Another specific use-case is when the raw data from different samples is not available. In this case, the researcher only has access to the parameter-values that are provided in the research papers. Because the `pbf()` function inherently converts objects to `bain` objects, each sample requires three pieces of information in order to be usable: the named parameter estimates, a covariance matrix of their standard errors and the sample size. First these values are extracted from all samples. A function is created that extracts this information which is then used to extract the necessary information from all samples.

```{r}
get_estimates <- function(sample){

  # number of groups in the 'site' variable
  number_sites <- length(levels(sesamesim$site))

  # sample size of each sample
  sample_size <- nrow(sample)

  # obtains mean and standard deviation of postnumb for each group
  site_estimates <- sapply(1:number_sites, function(i){
    rbind(
      mean(sample[sample['site'] == i, "postnumb"]),
      sd(sample[sample['site'] == i, "postnumb"])
      )
    })

  # get the estimates of the mean postnumb for each group and give them appropriate names
  est_names <- sprintf("site%d", 1:number_sites)
  site_means <- site_estimates[1,]
  names(site_means) <- est_names

  # obtain covariance matrix of the estimates, samples are assumed to be independent.
  covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
  dimnames(covmat) <- list(est_names, est_names)

  return(list(means = site_means, covmat = covmat, sample_size = sample_size))
}

@@

Then, each object is converted to a `bain` object using the extracted information and combined in a list.
```{r}
bain_sample1_ests <- bain(x = ests_sample1[['means']], 
                          Sigma = ests_sample1[['covmat']], 
                          n = ests_sample1[['sample_size']], 
                          hypothesis = hypotheses)
